Cost- and Deadline-constrained Scheduling of Scientific Workflow Ensembles in IaaS Clouds

### 2012-01-20 paper/paper.tex
Spelling corrections

The number of workflows completed is a simplest metric and allows distinguishing
which algorithm performs better. However, as the workflows have
different priorities and sizes, a less efficient algorithm may schedule smaller
low-priority workflows first, thus increasing the total number of workflows
complete. It may also happen that too conservative admission algoritm may reject
overestimate the costs of some large workflows, thus rejest some high-priority
ones which actually could fit into the schedule. Therefore we defined a user
preference score as:

@@

We run the simulations for the constraint values ranging between the
estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
ca. 50-150 (budget, deadline) points for each application. The runs were
repeated for Pareto-distributed and costant workflow sizes. From this result
space we selected the most interesting plots which give the good overview  of
the observed algorithm behavior.

@@

\caption{Summary table for equal size workflows whowing number of times each
algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
ensembles of equal size workflows for Montage (M), CyberShake (C), LIGO (L), 
Epigenomics (E), SIPHT (S).
\label{tab:num-dags-equal}}
\end{table}

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric $T_C$. These plots confirm our previous observations about the
algorithms performance, but they also allow to observe a more general property.
For a given budget, when the deadline is tightened the number of work done
steeply decreases. Assiging a shorter deadline means that more VMs need to be
provisioned to complete the work, but this leads to higher parallelism and thus
to the lower parallel efficiency. E.g. running 2 VMs for 10 hours will be more
efficient than running 10 VMs for 2 hours. Therefore, if deadline permits and
costs are important, parallelism should be avoided when possible.

These observations are further confirmed when examining the efffective resource
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increase of actual costs of computing. This increase reaches
$\sim$20\%, which means that if we need to run e.g. a SIPHT workflow ensemble
with a short deadline, $\sim$\$200 out of \$1000 is spent on idle resources.

Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We can see that the selection of deadline and cost can be formulated as
multi-criteria optimization problem and the curves presented in
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the number of work done, which direcly results in minimization of effective cost parameter for a given
deadline. Therefore, the obtained results can be used to assist the decison in
resource allocation planning, when both cost and time criteria are given not as
constraints but as objectives.

@@

In order to compare efficiency of algorithms, we counted the number of times
each algorithm achieved the best results for all tested applications, deadlines
and budgets. If two or more algorithms achieve the same highest results, all are
counted as highest. Table~\ref{tab:num-dags-pareto} shows these summary values
for Pareto-distributed workflows. These results depend on the metric: number of
workflows completed does not count the priorities or sizes of workflows, which
means that it happens more often that all the algorithms achieve the same (best)
result. However, it does not mean that all algorithms complete the same
workflows: that is why values in the tables (a), (b) and (c) are different.
Exponential score and total amount of computing time complete give the same
rankings, since in our test ensembles larger workflows were given higher
priorities. It can be observed that based on the sampled problem space the
workflow-aware algorithms achieve highest score in $\sim$2--3 times more cases
than the simple DPDS which does not take into account workflow structure or
size. In average, SPSS performs better in $\sim$20--50\% more times than
WA-DPDS. We have to note, that the sampled paramter space included the Montage
and CyberShake workflows consisting of many small tasks, for which cases SPSS
does not perform equally well.

@@

Fig.~\ref{fig:budgets} shows the same data for SIPHT ensemble from another
perspective. It helps answer the question: given a deadline, how many workflows
can be finished when we increse the budget. In Fig.~\ref{fig:budgets}(a) the
growth is super-linear, what results from the distribution of workflow sizes in
the ensemble (see Fig.~\ref{fig:ensemble-distribution}). For ensemble of equal
size workflows (see Fig.~\ref{fig:budgets}(b)) the growth is nearly linear. It
can be also observed that staitc algorithms give better results for smaller
budgets, while dynamic algorithms for larger ones.

@@

Another general observation is that although clouds provide such capabilities as
dynamic adding and removing of resources at runtime and high parallelization due
to ilusion of unlimited resources, it turns out that is recommended to avoid
using these capabilities when possible, in order to achieve high resource
utilization and cost effectiveness. This observation comes from the fact that
starting and stopping more VMs incurres a cost which can be amortized by running
each started VM as long as possible before shutdown. This effect was confirmed
in the case of SPSS algorithm for Montage and CyberShake workflows with small
budgets. Similarly, high degree of parallelization (large number of nodes) leads
also to less efficient resource utilization and in the case of clouds the drop
in parallel efficiency means also the increase of cost. This result, clearly
visible in Fig~\ref{fig:cost}, leads to multi-criteria decision problems and
related trade-offs. E.g. our simulation results can be used as hints to answer
the question how much time and money is required to actually complete all the
workflows from the given ansemble.

@@

Our current study reveals multiple paths for future work. One interesting
question would be to answer how the static and dynamic algorithms perform in the
cases of increased uncertainities and failures, that have important influence
on overall system performance~\cite{Sakellariou2010,Dongarra2007}. Second
important goal would be to extend our application and infrastructure model to
include various data storage options available on clouds. The experimental study
on that subject~\cite{Juve2010} suggests that data demands of scientific
workflows have a high impact not only on execution time but also on costs on
commercial clouds. Finally, we would like to address the heterogeneity of the
infrastructure, as multiple instance types and providers, including private
and community clouds will make the problem more complex and
challenging~\cite{Marshall2010,vockler11,Juve2010}.

>>>>

The number of workflows completed is a simplest metric and allows distinguishing
which algorithm performs better. However, as the workflows have
different priorities and sizes, a less efficient algorithm may schedule smaller
low-priority workflows first, thus increasing the total number of workflows
complete. It may also happen that too conservative admission algorithm may reject
overestimate the costs of some large workflows, thus reject some high-priority
ones which actually could fit into the schedule. Therefore we defined a user
preference score as:

@@

We run the simulations for the constraint values ranging between the
estimated limits as indicated in Section~\ref{sec:scenarios}, resulting in
ca. 50-150 (budget, deadline) points for each application. The runs were
repeated for Pareto-distributed and constant workflow sizes. From this result
space we selected the most interesting plots which give the good overview  of
the observed algorithm behavior.

@@

\caption{Summary table for equal size workflows showing number of times each
algorithm achieved the highest score over all budgets and deadlines. Data from a total of 525 runs on 
ensembles of equal size workflows for Montage (M), CyberShake (C), LIGO (L), 
Epigenomics (E), SIPHT (S).
\label{tab:num-dags-equal}}
\end{table}

Fig.~\ref{fig:total-time} presents the same results using the total computing
time metric $T_C$. These plots confirm our previous observations about the
algorithms performance, but they also allow to observe a more general property.
For a given budget, when the deadline is tightened the number of work done
steeply decreases. Assigning a shorter deadline means that more VMs need to be
provisioned to complete the work, but this leads to higher parallelism and thus
to the lower parallel efficiency. E.g. running 2 VMs for 10 hours will be more
efficient than running 10 VMs for 2 hours. Therefore, if deadline permits and
costs are important, parallelism should be avoided when possible.

These observations are further confirmed when examining the effective resource
cost, as shown in Fig.~\ref{fig:cost}. It can be seen that lower resource
utilization and lower parallel efficiency for short deadlines lead to
substantial increase of actual costs of computing. This increase reaches
$\sim$20\%, which means that if we need to run e.g. a SIPHT workflow ensemble
with a short deadline, $\sim$\$200 out of \$1000 is spent on idle resources.

Another general conclusion can be drawn from Fig.~\ref{fig:cost}: despite some
random fluctuations resulting mostly from the finite result space, all the
curves for all the applications and budgets have a similar shape. This shape
represents a trade-off curve between two conflicting objectives: time and cost.
We can see that the selection of deadline and cost can be formulated as
multi-criteria optimization problem and the curves presented in
Fig.~\ref{fig:cost} are approximations of the Pareto front (or Pareto set) of
solutions to this problem. In our case the solutions were achieved by maximizing
the number of work done, which directly results in minimization of effective cost parameter for a given
deadline. Therefore, the obtained results can be used to assist the decision in
resource allocation planning, when both cost and time criteria are given not as
constraints but as objectives.

@@

In order to compare efficiency of algorithms, we counted the number of times
each algorithm achieved the best results for all tested applications, deadlines
and budgets. If two or more algorithms achieve the same highest results, all are
counted as highest. Table~\ref{tab:num-dags-pareto} shows these summary values
for Pareto-distributed workflows. These results depend on the metric: number of
workflows completed does not count the priorities or sizes of workflows, which
means that it happens more often that all the algorithms achieve the same (best)
result. However, it does not mean that all algorithms complete the same
workflows: that is why values in the tables (a), (b) and (c) are different.
Exponential score and total amount of computing time complete give the same
rankings, since in our test ensembles larger workflows were given higher
priorities. It can be observed that based on the sampled problem space the
workflow-aware algorithms achieve highest score in $\sim$2--3 times more cases
than the simple DPDS which does not take into account workflow structure or
size. In average, SPSS performs better in $\sim$20--50\% more times than
WA-DPDS. We have to note, that the sampled parameter space included the Montage
and CyberShake workflows consisting of many small tasks, for which cases SPSS
does not perform equally well.

@@

Fig.~\ref{fig:budgets} shows the same data for SIPHT ensemble from another
perspective. It helps answer the question: given a deadline, how many workflows
can be finished when we increase the budget. In Fig.~\ref{fig:budgets}(a) the
growth is super-linear, what results from the distribution of workflow sizes in
the ensemble (see Fig.~\ref{fig:ensemble-distribution}). For ensemble of equal
size workflows (see Fig.~\ref{fig:budgets}(b)) the growth is nearly linear. It
can be also observed that static algorithms give better results for smaller
budgets, while dynamic algorithms for larger ones.

@@

Another general observation is that although clouds provide such capabilities as
dynamic adding and removing of resources at runtime and high parallelization due
to illusion of unlimited resources, it turns out that is recommended to avoid
using these capabilities when possible, in order to achieve high resource
utilization and cost effectiveness. This observation comes from the fact that
starting and stopping more VMs incurs a cost which can be amortized by running
each started VM as long as possible before shutdown. This effect was confirmed
in the case of SPSS algorithm for Montage and CyberShake workflows with small
budgets. Similarly, high degree of parallelization (large number of nodes) leads
also to less efficient resource utilization and in the case of clouds the drop
in parallel efficiency means also the increase of cost. This result, clearly
visible in Fig~\ref{fig:cost}, leads to multi-criteria decision problems and
related trade-offs. E.g. our simulation results can be used as hints to answer
the question how much time and money is required to actually complete all the
workflows from the given ensemble.

@@

Our current study reveals multiple paths for future work. One interesting
question would be to answer how the static and dynamic algorithms perform in the
cases of increased uncertainties and failures, that have important influence
on overall system performance~\cite{Sakellariou2010,Dongarra2007}. Second
important goal would be to extend our application and infrastructure model to
include various data storage options available on clouds. The experimental study
on that subject~\cite{Juve2010} suggests that data demands of scientific
workflows have a high impact not only on execution time but also on costs on
commercial clouds. Finally, we would like to address the heterogeneity of the
infrastructure, as multiple instance types and providers, including private
and community clouds will make the problem more complex and
challenging~\cite{Marshall2010,vockler11,Juve2010}.
